---
title: 'Statistical report: Cause of Death in Heart Failure'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 6
    fig_width: 9
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB4, https://kiheartfailure.github.io/shfdb4/, v 4.0.0. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Cause of death

Cause of death was defined from ULORSAK in the Cause of Death Register:
```{r cd}
dccode <- tibble(
  "Death" = levscd,
  "ICD10" = c(
    "I21|I22", "I2[0-5] excl I21|I22", "I110|I130|I132|I255|I420|I423|I425|I426|I427|I428|I429|I43|I50|J81|K761|R57", "I6[0-4]",
    "I0[5-8]|I3[4-9]|Q22|Q23[0-3]|Z95[2-4]", "I48",
    "Other I",
    "J0[0-6]|J09|J1[0-8]|J4[0-4]",
    "E1[0-4]",
    "E0[0-7]|E1[5-6]|E2|E3[0-5]E7|E8|E9",
    "N1[7-9]",
    "C",
    "A|B",
    "U071|U072|U08|U09|U10|B342|B972",
    "K",
    "G",
    "D[5-8]",
    "F0[0-4]",
    "R99",
    "All other"
  )
)
dccode <- dccode %>%
  mutate(
    ICD10 = str_replace_all(ICD10, fixed("|"), ","),
    ICD10 = str_replace_all(ICD10, fixed("["), ""),
    ICD10 = str_replace_all(ICD10, fixed("]"), "")
  )
default_kable(dccode, caption = "ICD-10 codes Cause of Death", scale_down = F)
```

\clearpage

## CHECK OTHER CAUSES OF DEATH

```{r, child = "./src/tabcd.Rmd"}

```

\clearpage

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Association between EF and CV/Non-CV Death

All analyses are performed for partly 1 year follow-up and partly 5 years follow-up. 

Censoring was performed at the specific time point (1 or 5 years), migration from Sweden, end of follow-up (2021-12-31) and, for the cox proportional hazard regression, death from other causes, whichever came first. 

For patients with possibility for complete follow-up until 1 and 5 years respectively the proportion of deaths were presented in pie charts. 
OBS! det är inte samma patienter i pie vid 1 år som i incidens/modeller vid 1 år (i det senare även personer som är censurerade inom 1 år). 

Incidence rate per 100 person-years was presented for the different causes of death. 

Time to death was presented with cumulative incidence curves using the Kaplan-Meier method. 

Cox proportional hazards regressions were used to evaluate the association 
between EF and time to death. Partly without adjustment and partly adjusted for age and sex. As a consistency analysis death was modelled using a sub-distributional hazards model [@fg] 
where death from other causes was treated as a competing event. 

\blandscape

```{r, child = "./src/pieplot.Rmd"}

```

\elandscape

\clearpage

```{r, child = "./src/incplot.Rmd"}

```

\clearpage
 
```{r, child = "./src/km.Rmd"}

```

\clearpage

OBS OBS! F & G modeller bara körda på delmängd av data då tar så lång tid!!!!!!

```{r, child = "./src/outtab.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/causeofdeath. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
