---
title: 'Statistical report: Cause of Death in Heart Failure'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 6
    fig_width: 9
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB4, https://kiheartfailure.github.io/shfdb4/, v 4.0.1. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Cause of death

Cause of death was defined from ULORSAK in the Cause of Death Register:
```{r cd}
dccode <- tibble(
  "Death" = levscd,
  "ICD10" = c(
    "I21|I22", "I2[0-5] excl I21|I22", "I2[6-8]",
    "I110|I130|I132|I255|I420|I423|I425|I426|I427|I428|I429|I43|I50|J81|K761|R570",
    "I6[0-4]",
    "I71|I72|I79[0-1]",
    "I70|I73|I74|I77|I78|I79[2-8]",
    "I0[5-8]|I3[4-9]|Q22|Q23[0-3]|Q23[5-9]",
    "I48",
    "I",
    "J",
    "E1[0-4]",
    "E0[0-7]|E1[5-6]|E2|E3[0-5]|E7|E8|E9",
    "N1[7-9]",
    "C",
    "U071|U072|U08|U09|U10|B342|B972",
    "A|B excl B342|B972",
    "K",
    "G",
    "D[5-8]",
    "F0[0-4]|R54",
    "M",
    "V|W|X[0-5]",
    "X[6-7]|X8[0-4]",
    "R99",
    "All other"
  )
)
dccode <- dccode %>%
  mutate(
    ICD10 = str_replace_all(ICD10, fixed("|"), ","),
    ICD10 = str_replace_all(ICD10, fixed("["), ""),
    ICD10 = str_replace_all(ICD10, fixed("]"), "")
  )
default_kable(dccode, caption = "ICD-10 codes Cause of Death", scale_down = F, font_size = 7)
write.xlsx(dccode, paste0("./output/tabs/CDcode_", Sys.Date(), ".xlsx"), rowNames = FALSE)
```

\clearpage 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Association between EF and CV/Non-CV Death

All analyses are performed for partly 1 year follow-up and partly 5 years follow-up. 

Censoring was performed at the specific time point (1 or 5 years), migration from Sweden, end of follow-up (2021-12-31) and, for the cox proportional hazard regression, death from other causes, whichever came first. 
Unknown cause of death is not included in neither CV nor non-CV.

For patients with possibility for complete follow-up until 1 and 5 years respectively the proportion of deaths were presented in pie charts. Note that the patients in the pie charts are therefore a subset of the patients in the incidence ratio bar charts and models. 

Incidence rate per 1000 person-years was presented for the different causes of death. 

Time to death was presented with cumulative incidence curves. 

Sub-distributional hazards models [@fg] where death from other causes was treated as a competing event
were used to evaluate the association 
between EF and time to death. Partly without adjustment and partly adjusted for age and sex. As a consistency analysis death was modelled using Cox proportional hazards regressions where death from other causes were censored. 

The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death_5y / 365.25), dig = 1),
                                             min = fn(min(sos_outtime_death_5y / 365.25), dig = 1),
                                             max = fn(max(sos_outtime_death_5y / 365.25), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r rsdata %>% summarise(sumpy = fn(sum(sos_outtime_death_5y) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

```{r deskstat}

deskdeath_ef <- rsdata %>%
  filter(sos_deathcause_cat2_5y != "Alive") %>%
  group_by(shf_ef_cat) %>%
  count(sos_deathcause_cat2_5y) %>%
  mutate(p = n / sum(n) * 100, 
         np = paste0(n, " (", fn(p, 1), "%)")) %>%
  ungroup()

deskdeath_tot <- rsdata %>%
  filter(sos_deathcause_cat2_5y != "Alive") %>%
  count(sos_deathcause_cat2_5y) %>%
  mutate(p = n / sum(n) * 100, 
         np = paste0(n, " (", fn(p, 1), "%)"), 
         shf_ef_cat = "Total")

deskdeath <- bind_rows(deskdeath_tot, deskdeath_ef) %>%
  select(-n, -p) %>%
  rename('Cause of death' = sos_deathcause_cat2_5y) %>%
  pivot_wider(values_from = np, names_from = shf_ef_cat)

default_kable(deskdeath, caption = sanitize_text("N (%) causes of death"), scale_down = F)
```

\blandscape

```{r, child = "./src/pieplot.Rmd"}

```

\elandscape

\clearpage

```{r, child = "./src/incplot.Rmd"}

```

\clearpage

```{r, child = "./src/outtab.Rmd"}

```

\clearpage
 
```{r, child = "./src/km.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/causeofdeath. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
