---
title: 'Statistical report: Cause of Death in Heart Failure'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 6
    fig_width: 9
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Utestående frågor

- Ska byta referens på EF???? (nu är referens rEF)

- Nu har med uppföljning tom 2 år och all tillgänglig tid. 2 år bra? Ska lägga till annan tidpunkt?

- Nu är alla som dör in-hospital exkluderade (för det är så default databasen är konstruerat). Lite extra jobb ta med även de som dör in-hospital men om värdefullt no probs.

- Bra variabler i tab 1?

# Data handling

## Data source

SHFDB4, https://kiheartfailure.github.io/shfdb4/, v 4.0.0. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = F)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

\clearpage

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Association between EF and CV/Non-CV Death

Long term follow-up: The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death / 365.25), dig = 1),
                                             min = fn(min(sos_outtime_death / 365.25), dig = 1),
                                             max = fn(max(sos_outtime_death / 365.25), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r rsdata %>% summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

Censoring was performed at migration from Sweden, end of follow-up (2021-12-31) and, for the cox proportional hazard regression, death from other causes, whichever came first. For the time specific models (at 2 years), censoring was also performed at 2 years, f this came fist. 

Incidence rate per 100 person-years was presented for the different causes of death. 

Time to death was presented with cumulative incidence curves using the Kaplan-Meier method. 

Cox proportional hazards regressions were used to evaluate the association 
between EF and time to death. Partly without adjustment and partly adjusted for age and sex. As a consistency analysis death was modelled using a sub-distributional hazards model [@fg] 
where death from other causes was treated as a competing event. 

```{r, child = "./src/incplot.Rmd"}

```

\clearpage
 
```{r, child = "./src/km.Rmd"}

```

\clearpage

OBS OBS! F & G modeller bara körda på delmängd av data då tar så lång tid!!!!!!

```{r, child = "./src/outtab.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/causeofdeath. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
